---
title: "squid diet metabarcoding"
author: "Kimberly Ledger"
date: "2023-07-18"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Taking a look at Charlotte's squid diet samples. 

Inputs: This code starts with the sample and taxon table that Charlotte created by matching ASVs to taxons and condensing the table. 

# Load libraries and data 

load libraries 
```{r}
library(tidyverse)
#library(dplyr)
library(ggplot2)
```

load sample/taxon table
```{r}
results_table <- read.csv("results_filtered_and_condensed_CS.csv")
```

reformat data table
```{r}
results_long <- results_table %>%
  pivot_longer(cols = c(2:16), names_to = "taxon", values_to = "reads")
head(results_long)
```

plot everything
```{r}
ggplot(results_long, aes(x=Sample_ID, y=reads, fill=taxon)) +
  geom_bar(stat = "identity") + 
    theme_bw() +
  labs(
    y = "number of sequencing reads",
    x = "sample ID",
    title = "squid taxon") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )
```

calculate read total and proportions per sample 
```{r}
results_long <- results_long %>%
  group_by(Sample_ID) %>%
  mutate(total = sum(reads)) %>%
  mutate(prop = reads/total)
```


remove samples with no reads from plot 
```{r}
results_long %>%
  filter(total != 0) %>%
  ggplot(aes(x=Sample_ID, y=reads, fill=taxon)) +
  geom_bar(stat = "identity") + 
    theme_bw() +
  labs(
    y = "number of sequencing reads",
    x = "sample ID",
    title = "squid taxon") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )
```

remove samples with less than 100 reads 
```{r}
results_long %>%
  filter(total > 100) %>%
  ggplot(aes(x=Sample_ID, y=reads, fill=taxon)) +
  geom_bar(stat = "identity") + 
    theme_bw() +
  labs(
    y = "number of sequencing reads",
    x = "sample ID",
    title = "squid taxon") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )
```

```{r}
results_long %>%
  filter(total > 100) %>%
  ggplot(aes(x=Sample_ID, y=prop, fill=taxon)) +
  geom_bar(stat = "identity") + 
    theme_bw() +
  labs(
    y = "proportion of sequencing reads",
    x = "sample ID",
    title = "squid taxon") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )
```

```{r}
results_long %>%
  filter(total > 100) %>%
  filter(str_starts(Sample_ID, "N")) %>%
  ggplot(aes(x=Sample_ID, y=reads, fill=taxon)) +
  geom_bar(stat = "identity") + 
    theme_bw() +
  labs(
    y = "number of sequencing reads",
    x = "sample ID",
    title = "negative samples") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )
```

```{r}
results_long %>%
  filter(total > 100) %>%
  filter(!str_starts(Sample_ID, "N")) %>%
  filter(Sample_ID != "sturgeon") %>%
  ggplot(aes(x=Sample_ID, y=reads, fill=taxon)) +
  geom_bar(stat = "identity") + 
    theme_bw() +
  labs(
    y = "number of sequencing reads",
    x = "sample ID",
    title = "squid taxon") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )
```
```{r}
results_long %>%
  filter(total > 100) %>%
  filter(!str_starts(Sample_ID, "N")) %>%
  filter(Sample_ID != "sturgeon") %>%
  ggplot(aes(x=Sample_ID, y=prop, fill=taxon)) +
  geom_bar(stat = "identity") + 
    theme_bw() +
  labs(
    y = "prop of sequencing reads",
    x = "sample ID",
    title = "squid taxon") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )
```

```{r}
results_long %>%
  filter(total > 100) %>%
  filter(!str_starts(Sample_ID, "N")) %>%
  filter(Sample_ID != "sturgeon") %>% 
  filter(reads > 0) %>%
  group_by(taxon) %>%
  summarise(n_squid = n(),
            total_reads = sum(reads)) %>% #  ,
            #mean_prop = mean(prop)) %>%
  arrange(desc(total_reads))
```




